### 2024-11-06 teaching/math240f24/draft.tex
lecture notes 5 nov

We conclude that $\sigma$ is a cycle.\slug
\bye

\medskip\boldlabel Connectedness.
We say a graph $G = (V,E)$ is {\it connected} if for all $u,v\in V$ there exists
a walk (and hence a path) from $u$ to $v$. All of the graphs we have
drawn so far have been connected. Here is a graph that is disconnected:
\midinsert
$$\epsfbox{disconnected.ps}$$
\endinsert
\noindent This picture is meant to depict one graph on $8$ vertices, not three
different graphs.

@@

\medskip\boldlabel\datestamp{19}{} Bipartite graphs.
A graph $G = (V,E)$ is called {\it bipartite} if there is a partition
$V = A\cup B$ of the vertex set (so $A\cap B = \emptyset$) called
the {\it bipartition} such that each edge has one endpoint in $A$ and one endpoint
in $B$. For example, hypercubes are bipartite.

@@

The calculation above shows that $K_{\lceil n/2 \rceil, \lfloor n/2 \rfloor}$
always has $n$ vertices and $\lfloor n^2/4\rfloor$ edges. It is the largest possible number
of edges such that Mantel's theorem does not apply; that is, from Mantel's theorem alone,
we do not know whether
there must be a triangle in $K_{\lceil n/2 \rceil, \lfloor n/2 \rfloor}$, but adding
a single edge to $K_{\lceil n/2 \rceil, \lfloor n/2 \rfloor}$, Mantel's theorem tells
us there must be a triangle in the resulting graph.

We will actually be able to prove that $K_{\lceil n/2 \rceil, \lfloor n/2 \rfloor}$ does not
contain any triangles. This means that the lower bound of $\lfloor n^2/4\rfloor$ in Mantel's theorem
is the best possible one.

>>>>

We conclude that $\sigma$ is a cycle.\slug

\medskip\boldlabel\datestamp{05}{xi} Connectedness.
We say a graph $G = (V,E)$ is {\it connected} if for all $u,v\in V$ there exists
a walk (and hence a path) from $u$ to $v$. All of the graphs we have
drawn so far have been connected. Here is a graph that is disconnected:
\midinsert
$$\epsfbox{disconnected.ps}$$
\endinsert
\noindent This picture is meant to depict one graph on $8$ vertices, not three
different graphs.

@@

\medskip\boldlabel Bipartite graphs.
A graph $G = (V,E)$ is called {\it bipartite} if there is a partition
$V = A\cup B$ of the vertex set (so $A\cap B = \emptyset$) called
the {\it bipartition} such that each edge has one endpoint in $A$ and one endpoint
in $B$. For example, hypercubes are bipartite.

@@

The calculation above shows that $K_{\lceil n/2 \rceil, \lfloor n/2 \rfloor}$
always has $n$ vertices and $\lfloor n^2/4\rfloor$ edges. It is the largest possible number
of edges such that Mantel's theorem does not apply. In other words, from Mantel's theorem alone,
we do not know whether
there must be a triangle in $K_{\lceil n/2 \rceil, \lfloor n/2 \rfloor}$, but adding
a single edge to $K_{\lceil n/2 \rceil, \lfloor n/2 \rfloor}$ results in a graph that must
have a triangle by Mantel's theorem.
\bye

Soon we will actually be able to prove that $K_{\lceil n/2 \rceil, \lfloor n/2 \rfloor}$ does not
contain any triangles. This means that the lower bound of $\lfloor n^2/4\rfloor$ in Mantel's theorem
is the best possible one.
